---
title: "Homework Tidyverse"
subtitle: "Hamid Semiyari"
format:
  html:
    embed-resources: true
author: "Zak Zerhouni"
date: today
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  fig.height = 3,
  fig.width = 6,
  fig.align = "center"
)

ggplot2::theme_set(ggplot2::theme_bw())
```

```{r}
#| message: false
#| warning: false
if(!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, gapminder, broom)
```

# Problem 1:

Copy paste and run the tribble given below.

```{r}
tribble( ~x,    ~y,    ~w,    ~z,
              210,   300,   220,   180,
              102,   100,   119,   187,
              176,   175,   188,   173,
              87,    95,   91,     94,
              202,   210,  234,    218,
              110,   122,  131,    128,
) -> dt
# Copied and pasted the tribble to run it.
tribble( ~x,    ~y,    ~w,    ~z,
              210,   300,   220,   180,
              102,   100,   119,   187,
              176,   175,   188,   173,
              87,    95,   91,     94,
              202,   210,  234,    218,
              110,   122,  131,    128,
)
```

## 1_a:

- Use and show a `map` function to find the "mean" of each column of the `dt` data table

```{r}
# The map function is used to generate a list of the mean of each column
dt |> 
  set_names(c("x.Mean", "y.Mean", "w.Mean", "z.Mean")) |> 
  map(mean) |> data.frame()
```

## 1_b:

- Use and show a `map` function to find the "standard deviation" of each column of the `dt` data table.

```{r}
# The function we map is changed to sd for the standard deviation
dt |> set_names(c("x.SD", "y.SD", "w.SD", "z.SD")) |> map(sd) |> data.frame()
```

## 1_c:

- Use and show a map function that will calculate the "square root" of each value of each column of the data table `dt`.

```{r}
# The function mapped is changed to sqrt
dt |> set_names(c("x.sqrt", "y.sqrt", "w.sqrt", "z.sqrt")) |> map(sqrt) |> data.frame()

```


## 1_d:

- Use R code to find the "mean", "max", "1st Quartile", "3rd Quartile", "Median", and "Min" for each column of the `dt` data table. (Hint: You do not have to use a map function)

```{r}
# The summary function is used to display the min, 1st quartile, median, 
# mean, 3rd quartile, and the max.
dt |> summary()
```


<!-- ----- End of Problem 1 ----- -->

# Problem 2:

Write a function that uses a for loop that, for each iteration, randomly draws 5 observations from an exponential distribution with "rate" parameter 1 (use `rexp()`) and calculates its "mean". It should do this 10,000 times. Choose an appropriate plot to plot the distribution of "means".

- Solution: We want to fill a vector with values. We should create this vector beforehand using the `vector()` function.

- We want to create a vector of size 10 and we want its values to be numeric.

- Then we want to randomly draw 5 observations from an exponential distribution with "rate" parameter 1 (use `rexp(5, rate = 1)`) and calculates its "mean"

- Now we want to create a vector that has length of 10 and its values are mean of this randomly selected exponential functions

- Now we want to draw the histogram. But we need to have a data frame. Let us define a data frame.

- Now we can draw a histogram

  - **Now, let's solve the problem all in one step**

## 2_a:

- Repeat part 1 by using a `map_*()` function.

## 2_b:

- Repeat part 1 by using the `replicate()` function.

## 2_c:

- Use a another for loop that will print out plots for sample sizes of 5, 10, and 20 observations (instead of just 5).

<!-- ----- End of Problem 2 ----- -->

# Problem 3:

- Use and show R coding to calculate the "standard deviation" for each variable of the data table `mtcars` using the “Special For Loop Method”.

**Note:**

- There are two methods. Please see a quick explanation for both methods in below:

  - Method 1:

```{r}
  multi_return <- function() {
    my_list <- list("color" = "red", "size" = 20, "shape" = "round")
    return(my_list) 
  }
```

```{r}
a <- multi_return()
a$color

a<- multi_return()
a$shape

a<- multi_return()
a$size
```

- Method 2. The "Special For Loop Method"

```{r}
#| eval: false
output <- vector("double", ncol(df))  # 1. output (use appropriate data frame instead of df)
for (i in seq_along(df)) {            # 2. sequence
  output[[i]] <- median(df[[i]])      # 3. body
}
output
```

- **Solution:** coding to calculate the "standard deviation" for each variable

<!-- ----- End of Problem 3 ----- -->

# Problem 4:

Generate a sequence of dates from **January 1, 2015** to **December 31, 2025**, spaced by every **two months**. Extract the **year**, **quarter**, and **ISO week number** for each date.

<!-- ----- End of Problem 4 ----- -->

# Problem 5

Given the following dates, compute the difference in **months** and **weeks** between each consecutive pair.

``` r
sample_dates <- c("2018-03-15", "2020-07-20", "2023-01-10", "2025-09-05")
```
<!-- ----- End of Problem 5 ----- -->

# Problem 6

Given a list of mixed date formats, use `map()` and `possibly()` from `purrr` to safely convert them to **Date** format and extract the **month name**.

``` r
date_strings <- list("2023-06-10", "2022/12/25", "15-Aug-2021", "InvalidDate")
```

<!-- ----- End of Problem 6 ----- -->