---
title: "Homework Tidyverse"
subtitle: "Hamid Semiyari"
format:
  html:
    embed-resources: true
author: "Zak Zerhouni"
date: today
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  fig.height = 3,
  fig.width = 6,
  fig.align = "center"
)

ggplot2::theme_set(ggplot2::theme_bw())
```

```{r}
#| message: false
#| warning: false
if(!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, gapminder, broom)
```

# Problem 1:

Copy paste and run the tribble given below.

```{r}
tribble( ~x,    ~y,    ~w,    ~z,
              210,   300,   220,   180,
              102,   100,   119,   187,
              176,   175,   188,   173,
              87,    95,   91,     94,
              202,   210,  234,    218,
              110,   122,  131,    128,
) -> dt
# Copied and pasted the tribble to run it.
tribble( ~x,    ~y,    ~w,    ~z,
              210,   300,   220,   180,
              102,   100,   119,   187,
              176,   175,   188,   173,
              87,    95,   91,     94,
              202,   210,  234,    218,
              110,   122,  131,    128,
)
```

## 1_a:

- Use and show a `map` function to find the "mean" of each column of the `dt` data table

```{r}
# The map function is used to generate a list of the mean of each column
dt |> 
  set_names(c("x.Mean", "y.Mean", "w.Mean", "z.Mean")) |> 
  map(mean) |> data.frame()
```

## 1_b:

- Use and show a `map` function to find the "standard deviation" of each column of the `dt` data table.

```{r}
# The function we map is changed to sd for the standard deviation
dt |> set_names(c("x.SD", "y.SD", "w.SD", "z.SD")) |> map(sd) |> data.frame()
```

## 1_c:

- Use and show a map function that will calculate the "square root" of each value of each column of the data table `dt`.

```{r}
# The function mapped is changed to sqrt
dt |> set_names(c("x.sqrt", "y.sqrt", "w.sqrt", "z.sqrt")) |> map(sqrt) |> data.frame()

```


## 1_d:

- Use R code to find the "mean", "max", "1st Quartile", "3rd Quartile", "Median", and "Min" for each column of the `dt` data table. (Hint: You do not have to use a map function)

```{r}
# The summary function is used to display the min, 1st quartile, median, 
# mean, 3rd quartile, and the max.
dt |> summary()
```


<!-- ----- End of Problem 1 ----- -->

# Problem 2:

Write a function that uses a for loop that, for each iteration, randomly draws 5 observations from an exponential distribution with "rate" parameter 1 (use `rexp()`) and calculates its "mean". It should do this 10,000 times. Choose an appropriate plot to plot the distribution of "means".

```{r}
# Set a seed for reproducibility
set.seed(4670329)

# Write the function to average out 5 random exponential draws and then
# append each average to a vector that is returned.
exponential_draw <- function(n) {
  # Initialize an empty vector
  vector_of_averages <- c()
  for(i in 1:n) {
    # compute the mean of the 5 ran. exp draws with rate=1
    mu <- mean(rexp(5,1))
    # The rbind function is used to append the most recent mean in the loop.
    vector_of_averages<- rbind(vector_of_averages,mu,deparse.level = 0)
  }
  return(vector_of_averages)
}

# Save a vector of 10,000 averages.
ten_thousand_draws <- exponential_draw(10000)

# Plot the vector as a histogram to get a view of the shape of the 
# distribution.
ggplot(data = data.frame(ten_thousand_draws), aes(x = ten_thousand_draws))+geom_histogram(bins = 100)+labs(title = "Histogram of 10,000 Exponential Averages", x = "Average of 5 Rand. Exp Draws", subtitle = "Sample generated using a for loop")
```

## 2_a:

- Repeat part 1 by using a `map_*()` function.

```{r}
# The same seed is used.
set.seed(4670329)
# A vector of 10000 is created and then mapped with map_dbl. This produces
# a sample of averages of 5 random exponential values generated.
ten_thousand_draws.map <- vector(mode="numeric", length = 10000) |>  
  map_dbl(\(x) mean(rexp(5,1)))
# An anonymous function is used inside of map_dbl to reassign the 
# value in the original vector.

#Plot the vector as a histogram to get a view of the shape of the 
# distribution.
ggplot(data = data.frame(ten_thousand_draws.map), aes(x = ten_thousand_draws))+geom_histogram(bins = 100)+labs(title = "Histogram of 10,000 Exponential Averages", x = "Average of 5 Rand. Exp Draws",subtitle="Sample generated with the map function")
```


## 2_b:

- Repeat part 1 by using the `replicate()` function.

```{r}
set.seed(4670329)
# Replicate is compact enough to fit directly into the ggplot object.
ggplot(data = data.frame(draws = replicate(10000,mean(rexp(5,1)))),
       aes(x = draws)
       )+
  geom_histogram(bins=100)+
  labs(title = "Histogram of 10,000 Exponential Averages", x = "Average of 5 Rand. Exp Draws",subtitle = "Sample generated using the replicate function")
```


## 2_c:

- Use a another for loop that will print out plots for sample sizes of 5, 10, and 20 observations (instead of just 5).

```{r}
set.seed(4670329)
# Iterate through a list of the desired sample sizes in the for loop
for(i in c(5,10,20)) {
  # Using the replicate method, we generate our 10000 sample by 
  # generating i random exponential draws and taking the mean
  p <- ggplot(data = data.frame(draws = replicate(10000,mean(rexp(i,1)))),
       aes(x = draws)
       )+
  geom_histogram(bins=100)+
  labs(title = "Histogram of 10,000 Exponential Averages", x = paste( "Average of ", i, " Rand. Exp Draws" ),subtitle = "Sample generated using the replicate function")
  # The paste function is used to change the number of draws in the x label.
  # Each graphic has to be printed to render from within the for loop.
  print(p)
}
```


<!-- ----- End of Problem 2 ----- -->

# Problem 3:

- Use and show R coding to calculate the "standard deviation" for each variable of the data table `mtcars` using the “Special For Loop Method”.

The "Special For Loop Method"

```{r}
# Create a function to use the "Special For Loop Method"
dataframe_sd <- function(df) {
  # Create a vector as long as the number of features in our data set
  for_loop_output <- vector("numeric", ncol(df))
  # Iterate the four loop across the numeric indices of the data set
  for (i in seq_along(df)) {
    # Replace the value in the vector with the output of sd
    for_loop_output[[i]] <- sd(df[[i]])
  }
  # Return the output of the values of the sd function as a vector,
  # while renaming the columns with the suffix .sd
  final_output <- for_loop_output |> 
         set_names(colnames(df) |> 
             map_chr(paste, ".", 
                     as.character(quote(sd)), collapse="") |> 
               map_chr(str_replace_all, " ", ""))
  return(final_output)
}

dataframe_sd(mtcars)
```

<!-- ----- End of Problem 3 ----- -->

# Problem 4:

Generate a sequence of dates from **January 1, 2015** to **December 31, 2025**, spaced by every **two months**. Extract the **year**, **quarter**, and **ISO week number** for each date.

<!-- ----- End of Problem 4 ----- -->

# Problem 5

Given the following dates, compute the difference in **months** and **weeks** between each consecutive pair.

``` r
sample_dates <- c("2018-03-15", "2020-07-20", "2023-01-10", "2025-09-05")
```
<!-- ----- End of Problem 5 ----- -->

# Problem 6

Given a list of mixed date formats, use `map()` and `possibly()` from `purrr` to safely convert them to **Date** format and extract the **month name**.

``` r
date_strings <- list("2023-06-10", "2022/12/25", "15-Aug-2021", "InvalidDate")
```

<!-- ----- End of Problem 6 ----- -->